The effect of linking assumptions and number of response options on inferred scalar implicature rate

### 2018-05-07 writeup.tex
Masoud completed his tasks for draft 1

Let's examine an example. Assume three response options (wrong, neither,
right). Assume further that a third of participants each gave each of
the three responses, i.e., the distributions of responses is 1/3, 1/3,
and 1/3. Under the Strong link, we infer that this task yielded an
implicature rate of 2/3. Under the Weak link, we infer that this task
yielded an implicature rate of 1/3. This is quite a drastic difference
if we are for instance interested in whether scalar implicatures are
inference defaults and we would like to interpret an implicature rate of
above an arbitrary threshold (e.g., 50\%) as evidence for such a claim.
Under the Strong link, we would conclude that scalar implicatures are
not defaults. Under the Weak link, we would conclude that they are.

In the experiment reported in the following section, we presented
participants with exactly this setup. Different groups of participants
were presented with different numbers of response options. We
categorized their responses according to the Weak and the Strong link
and tested whether number of response options and categorization scheme
leads to different conclusions about implicature rates.

@@

@@

200 participants were recruited using Amazon Mechanical Turk (binary=50,
ternary53, quaternary=43, quinary=54). No participant was excluded from
the final analysis.

\subsubsection{Procedure}\label{procedure}

The study was administered online through Amazon Mechanical Turk.
Participants were introduced to a set of cards with pictures of one or
two animals (Figure \ref{fig:stimuli}). They were told that a
blindfolded fictional character called Bob is going to guess what
animals are on the card. On each trial, participants saw a card as well
as a sentence representing Bob's guess. For example, they saw a card
with a cat on it and read the sentence \enquote{There is a cat on the
card.} The study ended after 24 trials. At the end participants
optionally provided demographic information. We also asked participants
if they had any prior training in logic. You can access and view on the
study's \href{https://github.com/thegricean/si-paradigms/}{online
repository}.

\subsubsection{Design and Materials}\label{design-and-materials}

@@

The study had two main manipulaitons within participants: the type of
card and the type of guess. There were two types of cards. Cards with
only one animal on them and cards with two animals. Animals were chosen
from the following set: cat, dog, and elephant There were three types of
guesses: simple (e.g. \emph{There is a cat}), conjunctive (e.g.
\emph{There is a cat and a dog}), and disjunctive (e.g. \emph{There is a
cat or a dog}). In each trial, the animal labels used in the guess and
the animal images on the card may have no overlap (e.g.~Image: cat,
Guess: \emph{There is an elephant}), a partial overlap (e.g.~Image: cat,
Guess: \emph{There is a cat or a dog}), or a total overlap (e.g.~Image:
cat and dog, Guess: \emph{There is a cat or a dog}). Crossing the number
of animals on the card, the type of guess, and the overlap between the
guess and the card results in 12 different possible trial types. We
chose 8 trial types (Figure \ref{fig:trials}), balancing the number of
one-animal vs.~two-animal cards, simple vs.~connective guesses, and
expected true vs.~false trials. Three trials were randomly selected from
each of the 8 trial-types, for a total of 24 trials. The order of these
24 trials was randomized as well.

@@

The study also had a between participant manipulation of the number of
response options in the forced choice task. Participants were randomly
assigned to one of four different conditions. The conditions differed
with respect to the number of response options: binary (wrong
vs.~right), ternary (wrong, neither, right), quaternary (wrong, kinda
wrong, kinda right, right), and quinary (wrong, kinda wrong, neither,
kinda right, right). We wanted to see if the number of response options
in the forced choice task would affect our estimate of the task's
\enquote{implicature rate}.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/trials-1.pdf}
\caption{\label{fig:trials}Trial types represented by example cards and
guesses.}
\end{figure}

@@

Here we present the proportion that participants chose the response
options in each of the 8 trial types of the binary, ternary, quaternary,
and quinary tasks. Figure \ref{fig:binaryPlot} shows the proportion of
\enquote{right} and \enquote{wrong} responses in the binary task.
Starting from the leftmost column, participants considered a guess
\enquote{wrong} if the guessed animal was not on the card. Moving to the
second column, participants considered a guess \enquote{right} if the
animal on the card was mentioned. However, if only one of the two
animals on the card was mentioned (exhaustive trials), 14\% of the times
participants considered the guess \enquote{wrong}. Moving to the third
column, if a conjunction of animals was guessed while only one animal
was on the card, participants considered the guess to be
\enquote{wrong}. If a conjunction of animals was guessed and both
animals were present on the card, all participants considered the guess
to be \enquote{right} as expected. Moving to the forth column, if a
disjunction of animals was guessed and only one of the animals was on
the card, participants considered the guess to be \enquote{right} almost
all the time. However, if both animals were present (scalar trials),
38\% of the times participants considered the guess to be
\enquote{wrong}.

@@

Comparing the response patterns in binary to quinary tasks (Figures
\ref{fig:binaryPlot} to \ref{fig:quinaryPlot}), we can observe that in
implicature trials, participants are less likely to choose the endpoints
of the scale (i.e. \enquote{wrong} and \enquote{right}) as they are
given more intermediate options. This raises the possibilty that the
strict \enquote{wrong/right} responses in the tasks with fewer options
(binary and ternary) were not accurately reflecting participant
judgments of implicature trials and pushed their judgments to the
extreme ends. More generally, the data suggest that participant
judgments in contexts that give rise to implicatures fall within the
intermrediate level of the response scale and forced choice tasks with
few options (especially binary tasks) may risk depicting judgments more
extreme than they actually are.

@@

Comparing the exhaustive and scalar columns of Figure
\ref{fig:implicatureRatePlot}, we see that with a strong linking
assumption, there are slightly higher rates for scalar implicatures in
the binary and ternary tasks. With a weak linking assumption, there may
be slightly higher rates for scalar implicatures in the binary and
ternary while the rates may be lower in the quaternary and quinary
tasks. In what follows, we test the effect of linking assumption and
response options on exhaustive and scalar implicature rates more
formally.

@@

For our formal analysis, we use a binomial mixed effects model with the
fixed effect of response type (binary, ternary, quaternary, quinary),
linking assumption (strong vs.~weak), and trial type (exhaustive
vs.~scalar), as wellas the random intercept for participants as well as
random intercept and slope for items (the card they saw).

All the analytical decisions described here were pre-registered and you
can access them through the following link:
\url{https://aspredicted.org/tq3sz.pdf}

\begin{itemize}
\tightlist
\item
  Was logical training a significant predictor of \enquote{implicature
  rate}?
\end{itemize}

\section{\# Discussion}\label{discussion}

@@

@@

>>>>

Let's examine an example. Assume three response options (wrong, neither,
right). Assume further that a third of participants each gave each of
the three responses, i.e., the distributions of responses is 1/3, 1/3,
and 1/3. Under the Strong link, we infer that this task yielded an
implicature rate of 2/3. Under the Weak link, we infer that this task
yielded an implicature rate of 1/3. This is quite a drastic difference
if we are for instance interested in whether scalar implicatures are
inference defaults and we would like to interpret an implicature rate of
above an arbitrary threshold (e.g., 50\%) as evidence for such a claim.
Under the Strong link, we would conclude that scalar implicatures are
not defaults. Under the Weak link, we would conclude that they are. In
the experiment reported in the following section, we presented
participants with exactly this setup.

@@

In this study, we presented participants with an online card game.
Different groups of participants were presented with different numbers
of response options for the task. In critical trials, participants were
presented with descriptions for the cards that typically result in
exhaustive or scalar implicatures. We categorized their responses in
such trials according to the Weak and the Strong link, and tested
whether the number of response options and the linking assumptions lead
to different conclusions about the rate of computed implicatures in the
experimental task.

@@

200 participants were recruited using Amazon Mechanical Turk. They
optionally provided demographic information at the end of the study.
Mean age of the participnats was 35. We also asked participants if they
had any prior training in logic. 40 participants reported that they had
while 160 had no prior training in logic. No participant was excluded
from the final analysis.

\subsubsection{Materials and Procedure}\label{materials-and-procedure}

The study was administered online through Amazon Mechanical Turk.
Participants were first introduced to the set of cards we used in the
study (Figure \ref{fig:stimuli}). Each card had pictures of one or two
animals. Animals were chosen from the following set: cat, dog, and
elephant. Then participants were introduced to a blindfolded fictional
character called Bob. Bob was blindfolded to control for violations of
ignorance expectations with disjunction. Participants were told that Bob
is going to guess what animals are on the card. We asked participants to
let us know whether Bob's guess is wrong or right. In each trial,
participants saw a card as well as a sentence representing Bob's guess.
For example, they saw a card with a cat and read the sentence
\enquote{There is a cat on the card.} Depending on the task participants
were assigned to, they had to choose between two (binary task), three
(ternary task), four (quaternary task), or five (quinary task) response
options. The study ended after 24 trials. You can access and view the
study using the paper's
\href{https://github.com/thegricean/si-paradigms/}{online repository}.

\subsubsection{Design}\label{design}

@@

The study had two main manipulaitons within participants: the type of
card and the type of guess. There were two types of cards. Cards with
only one animal on them and cards with two animals. There were three
types of guesses: simple (e.g. \emph{There is a cat}), conjunctive (e.g.
\emph{There is a cat and a dog}), and disjunctive (e.g. \emph{There is a
cat or a dog}). In each trial, the animal labels used in the guess and
the animal images on the card may have no overlap (e.g.~Image: cat,
Guess: \emph{There is an elephant}), a partial overlap (e.g.~Image: cat,
Guess: \emph{There is a cat or a dog}), or a total overlap (e.g.~Image:
cat and dog, Guess: \emph{There is a cat or a dog}). Crossing the number
of animals on the card, the type of guess, and the overlap between the
guess and the card results in 12 different possible trial types. We
chose 8 trial types, balancing the number of one-animal vs.~two-animal
cards, simple vs.~connective guesses, and expected true vs.~false
trials. Three trials were randomly selected from each of the 8
trial-types, for a total of 24 trials. The order of these 24 trials was
randomized as well.

@@

The study also had a between participant manipulation of the number of
response options in the forced choice task. Participants were randomly
assigned to one of four different conditions or tasks. The tasks
differed with respect to the number of response options: binary (wrong
vs.~right), ternary (wrong, neither, right), quaternary (wrong, kinda
wrong, kinda right, right), and quinary (wrong, kinda wrong, neither,
kinda right, right). We wanted to see if the number of response options
in the forced choice task would affect our estimate of the task's
\enquote{implicature rate}.

@@

The experiment had 50 participants in the binary task, 53 in the ternary
task, 43 in the quaternary task and 54 in the quinary task. In this
section, we present the proportion that participants chose different
response options in each of the 8 trial types of these four tasks.

Figure \ref{fig:binaryPlot} shows the proportion of \enquote{right} and
\enquote{wrong} responses in the binary task. Starting from the leftmost
column, participants considered a guess \enquote{wrong} if the guessed
animal was not on the card. Moving to the second column, participants
considered a guess \enquote{right} if the animal on the card was
mentioned. However, if only one of the two animals on the card was
mentioned (exhaustive trials), 14\% of the times participants considered
the guess \enquote{wrong}. Moving to the third column, if a conjunction
of animals was guessed while only one animal was on the card,
participants considered the guess to be \enquote{wrong}. If a
conjunction of animals was guessed and both animals were present on the
card, all participants considered the guess to be \enquote{right} as
expected. Moving to the forth column, if a disjunction of animals was
guessed and only one of the animals was on the card, participants
considered the guess to be \enquote{right} almost all the time. However,
if both animals were present (scalar trials), 38\% of the times
participants considered the guess to be \enquote{wrong}.

@@

Comparing the response patterns in binary to quinary tasks (Figures
\ref{fig:binaryPlot}-\ref{fig:quinaryPlot}), we observe that in
implicature trials, participants are less likely to choose the endpoints
of the scale (i.e. \enquote{wrong} and \enquote{right}) as they are
given more intermediate options.

semantic violations vs.~pragmatic violations

@@

Comparing the exhaustive and scalar columns of Figure
\ref{fig:implicatureRatePlot}, we see that with a strong linking
assumption, there are slightly higher rates for scalar implicatures in
the binary and ternary tasks. With a weak linking assumption, there may
be slightly higher rates for scalar implicatures in the binary and
ternary while the rates may be lower in the quaternary and quinary
tasks. In what follows, we formally test the effect of linking
assumption and response options on exhaustive and scalar implicature
rates.

@@

\begin{table}

\caption{\label{tab:implicatureRate}\label{tab:modeltable}Model Parameter Estimates and Their Credible Intervals}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Predictors & Estimate & 2.5\% & 97.5\%\\
\midrule
Intercept & -8.60 & -13.98 & -4.53\\
Link = Weak & -0.15 & -4.86 & 4.77\\
Task = Quaternary & -1.83 & -8.08 & 4.20\\
Task = Quinary & -4.05 & -10.90 & 2.38\\
Task = Ternary & -1.45 & -7.31 & 4.56\\
\addlinespace
Implicature = Scalar & 6.09 & 1.00 & 12.29\\
Link = Weak : Task = Quaternary & 14.03 & 7.24 & 21.88\\
Link = Weak : Task = Quinary & 17.28 & 10.64 & 25.80\\
Link = Weak : Task = Ternary & 3.81 & -1.49 & 9.22\\
Link = Weak : Implicature = Scalar & 0.90 & -4.01 & 6.43\\
\addlinespace
Task = Quaternary : Implicature = Scalar & -5.67 & -13.66 & 1.54\\
Task = Quinary : Implicature = Scalar & -2.31 & -9.30 & 4.61\\
Task = Ternary : Implicature = Scalar & -1.31 & -7.70 & 4.65\\
Link=Weak : Task=Quaternary : Implicature=Scalar & -3.29 & -12.07 & 4.55\\
Link=Weak : Task=Quinary : Implicature=Scalar & -7.74 & -16.59 & -0.16\\
Link=Weak : Task=Ternary : Implicature=Scalar & -1.44 & -7.00 & 4.22\\
\bottomrule
\end{tabular}
\end{table}

For our formal analysis, we used a bayesian binomial mixed effects model
using the R packge \enquote{brms} (Bürkner \& others, 2016). The model
had the fixed effects of response type (binary, ternary, quaternary,
quinary), linking assumption (strong vs.~weak), and trial type
(exhaustive vs.~scalar), as well as their two way and three way
interactions. We also included random intercepts and slopes for items
(cards) and participants. Following Barr, Levy, Scheepers, and Tily
(2013), we used the maximal random effects structure by including random
slopes for all our fixed effects and their interactions. Since the
number of response options was a between participant variable we did not
include random slopes of response options for participants. Four chains
converged after 2000 iterations each (warmup = 1000). Table
\ref{tab:modeltable} summarizes the mean parameter estimates and their
95\% credible intervals. \(\hat{R}=1\) for all estimated parameters. All
the analytical decisions described here were pre-registered\footnote{You
  can access our pre-registration at
  \url{https://aspredicted.org/tq3sz.pdf}}.

The model provided evidence for two effects in the study. First, there
was a main effect of trials such that scalar trials had slightly higher
implicature rates than exhaustive trials (Mean Estimate = 6.09, 95\%
Credible Interval={[}1, 12.29{]}). Second, there was an interaction
between linking assumption and number of resopnse options such that the
quaternary task (Mean Estimate = 14.03, 95\% Credible Interval={[}7.24,
21.88{]}) and the quinary task (Mean Estimate = 17.28, 95\% Credible
Interval={[}10.64, 25.80{]}) with a weak linking assumption had higher
implicature rates.

\subsection{Discussion}\label{discussion}

We asked whether the linking assumptions and the number of response
options available to participants affects the inferred implicature rate
in an experimental study. The results presented here suggest they do. A
linking assupmtion that considered the highest point on the scale as
literal and any lower point as pragmatic (weak link) reported higher
implicature rates in tasks with 4 or 5 options. A linking assumption
that considered the lowest point on the scale as pragmatic and any
higher point as literal (strong link) reported lower implicature rates
in tasks with 4 or 5 options. The results suggest that the choice of
linking assumption is a crucial analytical step that can significantly
impact the conclusions of truth value judgment tasks with more than two
response options. The lower rate of implicatures with a strong linking
assumption implies that in such studies, strong linking assumptions may
underestimate participants' pragmatic competence.

While the binary truth value judgement task avoids the analytic decision
between strong and weak linking assumptions, our results suggest that
binary tasks can also underestimate participants' pragmatic competence.
In binary tasks, participants are often given the lowest and highest
points on a scale (\enquote{wrong} vs. \enquote{right}) and are asked to
report pragmatic infelicities using the lowest point (e.g.
\enquote{wrong}). Our study showed that in trials with true but
infelicitous descriptions (implicature trials), participants often
avoided the lowest point on the scale if they were given more
intermediate options. Even though the option \enquote{wrong} was
available to participants in all tasks, participants in tasks with
intermediate options chose it less often. In computing implicature rate,
this pattern manifested itself as a decrease in implicature rate with
strong link when more response options were provided, and increase in
implicature rate with weak link when more response options were
provided. These conclusions are in line with Katsos and Bishop
(2011)\enquote{s argument that pragmatic violations are not as severe as
semantic violations and participants do not penalize them as much.
Providing participants with only the extreme opposits of the scale
(e.g.~wrong/right, false/true) when pragmatic violations are considered
to be of an intermediate nature risks misrepresentation of participants}
pragmatic competence.

This study did not investigate the effect of option labels on the
inferred implicature rate. However, the results provided suggestive
evidence that some options better capture participant intuitions of
pragmatic infelicities than others. Among the intermediate options,
\enquote{kinda right} was chosen most often to report pragmatic
infelicities. The option \enquote{neither} was rarely used in the
ternary and quinary tasks (where it was used as a midpoint), suggesting
that participants found pragmatic infelicities as different degrees of
being \enquote{right} and not \enquote{neither right nor wrong.}
Therefore, options that capture grades of being \enquote{right} like
\enquote{kinda right} proved be most suitable for capturing true but
infelicitous utterances.

This study had three design features that we would like to investigate
in future work. First, the utterances were by a blindfolded character
because we wanted to control for violation of ignorance expectations
with disjunction. A disjunction such as \enquote{A or B} often carries
an implication or expectation that the speaker is not certain which
alternative actually holds. In future work, we would like to see how the
violation of the ignorance expectation would affect the inferred
implicature rate by having the fictional character describe the cards
while looking at them. Second, in this study we considered exhaustive
and scalar implicatures with \emph{or}. We would like to see if similar
effects hold for the scalar implicatures with \emph{some}. Finally, our
experiment was designed as a guessing game and the exact goal of the
game was left implicit. We expect that different goals, for example help
the character win more points vs.~help the character be more accurate,
would affect how strict or lenient participants are with their judgments
and ultimately affect the implicature rate in the task. In future work
we would like to systematically vary the goal of the game and explore
its effects on the inferred implicature rate.

\section{General Discussion}\label{general-discussion}

@@

\hypertarget{ref-barr2013random}{}
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random
effects structure for confirmatory hypothesis testing: Keep it maximal.
\emph{Journal of Memory and Language}, \emph{68}(3), 255--278.

@@

\hypertarget{ref-burkner2016brms}{}
Bürkner, P.-C., \& others. (2016). Brms: An r package for bayesian
multilevel models using stan. \emph{Journal of Statistical Software},
\emph{80}(1), 1--28.
